---
title             : "Investigating the Interaction of Direct and Indirect Relation on Memory Judgments and Retrieval"
shorttitle        : "Judgments and Recall"

author: 
  - name          : "Nicholas P. Maxwell"
    affiliation   : "1,"
  - name          : "Erin M. Buchanan"
    affiliation   : "2"
    corresponding : yes    # Define only one corresponding author
    address       : "326 Market St., Harrisburg, PA, 17101"
    email         : "ebuchanan@harrisburgu.edu"

affiliation:
  - id            : "1"
    institution   : "University of Southern Mississippi"
  - id            : "2"
    institution   : "Harrisburg University of Science and Technology"

author_note: >
  Nicholas P. Maxwell is a Ph.D. candidate at the University of Southern Mississippi. Erin M. Buchanan is an Professor of Cognitive Analytics at Harrisburg University of Science and Technology. 
  
abstract: >
  This study examined the interactive relationship between direct and indirect relation strength in the prediction of item judgments and cued-recall performance. Participants were recruited from Amazon's Mechanical Turk and were given word pairs of varying relatedness to judge for their semantic, thematic, and associative strength. After completing a distractor task, participants then completed a cued recall task. First, we sought to expand previous work on judgments of associative memory (JAM) to include semantic and thematic based judgments, while also replicating bias and sensitivity findings. Next, we tested for an interaction between direct and indirect association when predicting participant judgments while also expanding upon previous work by examining that interaction when predicting recall. The interaction between direct and indirect association was significant for both judgments and recall. For low indirect association, direct association was the primary predictor of both judgment strength and recall proportions. However, this trend reversed for high indirect association, as higher levels of indirect relation decreased the effectiveness of direct relation as a predictor. Overall, our findings indicate the degree to which the processing of associative, semantic, and thematic information impacts cognitive processes such as retrieval and item judgments, while also examining the underlying, interactive relationship that exists in language used to represent concept information.

keywords          : "judgments, memory, association, semantics, thematics"

bibliography      : ["nick_ref.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
replace_ampersands: yes
csl               : apa6.csl
---

```{r libraries, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)

p.value = function(x){
  if (x < .001) { return("< .001")}
  else { return(apa(x, 3, F))}
}

pasteit = function(M, SD){
  return(paste(printnum(M, gt1 = F), " (",
        printnum(SD, gt1 = F), ")", sep = ""))
}

library(papaja)
library(cowplot)
library(MOTE)
library(ggplot2)
library(ez)
library(lsa)
library(kableExtra)

cleanup = theme(panel.grid.major = element_blank(), 
                panel.grid.minor = element_blank(), 
                panel.background = element_blank(), 
                axis.line.x = element_line(color = "black"),
                axis.line.y = element_line(color = "black"),
                legend.key = element_rect(fill = "white"),
                text = element_text(size = 15))
```

The study of cognition has a rich history of exploring the way in which associations affect human memory. One key finding is that associations between items influence cognitive processing and play a critical role in how well an individual retains learned information. Throughout the mid-20th century, researchers investigated this notion, particularly through the use of paired-associate learning (PAL). In this paradigm, participants are presented with a pair of items and are asked to make connections between them so that the presentation of one item (the cue) will in turn trigger the recall of the other (the target). Early studies of this nature focused primarily on the effects of meaning and imagery on recall performance. For example, @Smythe1968 found that noun imagery played a crucial role in PAL performance; subjects were much more likely to remember word-pairs that were low in meaning similarity if imagery between the two was high. Subsequent studies in this area focused on the effects of mediating variables on PAL tasks as well as the effects of imagery and meaningfulness on associative learning [@Richardson1998], with modern studies shifting their focus towards a broad range of applied topics such as how PAL is affected by aging [@Hertzog2002], its impacts on second language acquisition [@Chow2014], and even in evolutionary psychology [@Schwartz2013]. 

Early PAL studies routinely relied on stimuli generated from word lists that focused extensively on measures of word frequency, concreteness, meaningfulness, and imagery [@Paivio1969]. However, the word pairs in these lists were typically created due to their apparent relatedness or frequency of occurrence in text. While lab self-generation appears face valid, one finds that this method of selection lacks a decisive method of defining the underlying relationships between the pairs [@Buchanan2010], as these variables only capture psycholinguistic measurements of an individual concept (i.e., how concrete is *cat* and word occurrence). PAL is, by definition, used on word pairs, which requires examining concept relations in a reliable manner. As a result, free association norms have become a common means of indexing associative strength between word pairs [@Nelson2000]. 

## Measuring Association

Within cognitive psychology, word associations have been conceptualized differently across various lines of research (i.e., direct word associations, mediated associates, etc.; see @DeDeyne2013 for a review).  For the present study, we focus on two types of associations: direct associations and indirect associations. Direct word associations are traditionally viewed as the context-based relation between concepts, usually found in text or popular culture [@Nelson2000]. Within this framework, word associations are thought to arise in several different ways. Such associations may develop through their co-occurrence together in either written or spoken language. The terms *peanut* and *butter* have become associated over time through their joint use to depict a particular type of food, though separately, the two concepts share very little overlap in terms of meaning. However, this separation is not the case for all associative pairs. For example, word associations capture the knowledge that fish live in water (e.g., *fish* – *swim*) and that dogs and cats share many similar features. To generate norms measuring direct associations, participants engage in a free association task, in which they are presented with a cue word and are asked to list the first related target word that comes to mind. The probability of producing a given response to a particular cue word (i.e., the pair’s normed forward strength, FSG) can then be determined by dividing the number of participants who produced the response in question by the total number of responses generated for that word [@Nelson2000]. Thus, the free association process can be thought of as generating an index that contains the relative accessibility of related words in memory [@Nelson2004]. 

Using this technique, researchers have developed databases of associative word norms that can be used to generate stimuli, generally with a high degree of reliability [e.g., The University of South Florida Free Association Norms; @Nelson2004]. However, this reliability becomes questionable for weak associates. Because the traditional free association task focuses only the first word that is provided the cue, target items that are more weakly associated may become underrepresented in the dataset, as the inclination to respond with stronger associates may disrupt access to weaker associates (i.e., the availability heuristic]. Recently, The Small World of Words project [SWOW, @DeDeyne2013; @DeDeyne2018] has sought to correct for this sampling issue by employing a multiple response free association task. In this modified free association task, subjects are asked to generate three target items in response to the cue. The updated SWOW association norms provide several advantages when compared to other collections of free association norms. First, this norm set is the largest to date, consisting of approximately 12,000 cue items (for comparison, the USF norms consist of 5,400 cue items). Because of its large size, the SWOW norms provide a better approximation of natural language. Second, the use of a multiple response technique allows for greater reliability of weak associates, resulting in more weak associations being captured by the network, as weak associates are rarely given as the first response and thus may be under represented when only one response is elicited [@DeDeyne2013]. 

## Measuring Relatedness

Whereas direct associations focus on the relationships between individual words, indirect associations focus on how a concept fits into the overall structure of the semantic network [@DeDeyne2013; @Deese1965]. Because indirect associations capture information derived from the overall structure of the semantic network, these norms can also be used to represent semantic properties of item pairs and can be used to approximate links between concepts within semantic memory networks. This includes mediated associates [i.e., *lion* – *stripes* is mediated through *tiger*; see @Huff2011 for a review of mediated associates] and is one of the underlying factors behind distributional models of semantic memory [e.g., Latent Semantic Analysis, @Landauer199; Hyperspace Analogue to Language Model, @Lund1996]. These models posit that semantic representations are created through the co-occurrences of words together within a body of text and suggest that words with similar meanings will appear together in similar contexts [@Riordan2011]. On the other hand, connectionist models of semantic memory [e.g., @Rogers2006; @Rumelhart1986] portray the semantic network as a system of interconnected units representing concepts, which are linked together by weighted connections representing knowledge. By triggering the input units, activation will then spread throughout the system activating or suppressing connected units based on the weighted strength of the corresponding unit connections [@Jones2015]. 

Semantic overlap between concepts can measured in several ways. Feature production tasks [@McRae2005; @Vinson2008; @Buchanan2013] provide one means of generating semantic word norms. In such tasks, participants are shown the name of a concept and are asked to list what they believe the concept's most important features to be [@McRae2005]. Several statistical measures have been developed which measure the degree of feature overlap between concepts. First, similarity between any two concepts can be measured by representing them as vectors and calculating the cosine value (COS) between them [@Maki2004], with COS values ranging from 0 (completely unrelated) to 1 (perfectly related). For example, the pair *hornet* - *wasp* has a COS of .88, indicating a high degree of overlap between the two concepts. Feature overlap can also be measured by JCN, which involves calculating both the information content value of each concept and the lowest super-ordinate shared by each concept using an online dictionary, such as WordNET [@Miller1995]. The JCN value is then computed by summing together the difference of each concept and its lowest super-ordinate [@Jiang1997; @Maki2004]. The advantage to using COS values over JCN values is the limitation of JCN being tied to a somewhat static dictionary database, while a semantic feature production task can be used on any concept to calculate COS values. However, JCN values are less time consuming to obtain if both concepts are in the database [@Buchanan2013]. 

Finally, indirect associations computed from a large dataset can also be used as a measure semantic overlap, and indeed may provide a better measure semantic relatedness relative to feature production norms. @DeDeyne 2013 constructed a semantic network based on the distributions of associations (e.g., indirect associates) by converting free association data taken from the SWOW project into a weighted semantic network.  Computing the cosine overlap between the distribution of free association answers on any two concepts within this network provides a useful measure of meaning. Discussion of these measures then leads to the question of whether each one is truly assessing some unique concept or if they simply tap into various elements of our overall linguistic knowledge. Previous clustering and factor analyses  by @Maki2008 indicates that there are potentially three separate latent structures represented by these various measures of similarity – associative, semantic, and thematic types of relatedness. However, another interpretation of their results is that the data collection of the measurement matters – variables that are based on participant responses to cued stimuli grouped together, while text-corpora based and WordNET based similarity measures separated into separate factors. By using the participant responses from SWOW to measure indirect association,  we draw from a larger, newer set of data and resolve a potential confound of conflating measurement techniques. 

## Application to Judgment Studies

Traditional judgment of learning tasks (JOL) can be viewed as an application of the PAL paradigm; participants are given pairs of items and are asked to judge how accurately they would be able to correctly respond with the target with the cue on a recall task. Judgments are typically made out of 100, with a participant response of 100 indicating full confidence in recall ability. In their 2005 study, Koriat and Bjork examined overconfidence in JOLs by manipulating associative relations (forward strength from @Nelson2004) between word-pairs and found that subjects were more likely to overestimate recall for pairs with little or no associative relatedness. Additionally, this study found that when accounting for associative direction, subjects were more likely to overestimate recall for pairs that were high in backwards strength (i.e., the likelihood of the target when shown the cue) but low in forward strength. To account for this finding, the authors suggested that JOLs may rely more heavily on overlap between cue and target with the direction of the associative relationship being secondary. For example, the pair *bird* - *feather* in the SWOW norms appears to have a low forward strength (.031) and a higher backward strength (.199). However,  the indirect relation between *bird* and *feather is .063. Therefore, it is important to investigate what may lead to the perceived relatedness between the item pairs and resulting in inflated JOLs.

The JOL task can then be manipulated to investigate perceptions of word pair relation by having participants judge how related they believe the cue and target items to be [@Maki2007a; @Maki2007]. The judged values generated from this task can then be compared to the normed databases to create a similar accuracy function or correlation as is created in JOL studies. When presented with the item pair, participants are asked to estimate the number of people out of 100 who would provide the target word when shown only the cue [@Maki2007], which mimics how association word norms are created through free association tasks. @Maki2007a investigated such judgments within the context of associative memory by having participants rate how much associative overlap was shared between normed item pairs and found that responses were greatly overestimated relative to the actual normed overlap strength for pairs that were weak associates, while underestimated for strong associates, thus replicating the @Koriat2005 findings for relatedness judgments based upon associative memory, rather than judgments based on learning.

The judgment of associative memory (JAM) function provides one means of visualizing the influence various cognitive biases have on associative memory judgments. By plotting the judged values against the word pair's normed associative strength, a fit line can be calculated which displays the calibration of JAM ratings relative to normed associative strength. When plotted, these judgments characteristically have a high intercept (an overestimation bias) along with a shallow slope (low sensitivity to changes in normed pair strength). Figure \@ref(fig:makislope) illustrates this function. Overall, the JAM function has been shown to be highly reliable and generalizes well across multiple variations of the study, with item characteristics such as word frequency, cue set size (QSS), and semantic similarity all having a minimal influence on it [@Maki2007]. Furthermore, an applied meta-analysis of more than ten studies on JAM indicated that bias and sensitivity are nearly unchangeable, often hovering around 40-60 points for the intercept and .20-.30 for the slope [@Valentine2013]. Additionally, the @Valentine2013 study extended this research to include judgments of semantic memory with the same results. Finally, @DeDeyne2013a found that JAM ratings for weak and moderate associates are best predicted by continuous response association norms relative to traditional free association norms.

```{r makislope, echo=FALSE, fig.cap = "JAM slope findings from Maki (2007a). JAM is characterized by a high intercept (between 40 and 60) and a shallow slope (between 0.20 and 0.40). The solid line shows expected results if judgment ratings are perfectly calibrated with association norms.", fig.height=6, fig.width=6}

####JAM plot####
fakedata = data.frame(FSG = 0:100,
                      JAM = 0:100)

plot1 = ggplot(fakedata, aes(FSG, JAM)) + 
  xlab("Forward Strength") +
  ylab("Participant Judgments") +
  geom_abline(aes(intercept = 0, slope = 1, linetype = "Expected")) +
  geom_abline(aes(intercept = 50, slope = .30, linetype = "JAM")) + 
  scale_size_continuous(guide = FALSE) +
  scale_linetype_manual(values = c("solid", "dashed"),
                        breaks = c("Expected", "JAM"),
                        name = "Function") +
  coord_cartesian(xlim = c(0,100), ylim = c(0,100)) +
  geom_vline(xintercept = -5) +
  geom_hline(yintercept = -5) +
  cleanup 
  
plot1 + theme(legend.position = c(.75, .25))

```

The discrepancy between direct association strength and JAM ratings is noteworthy because on the surface, the two tasks should each be tapping into the same concept of associative overlap. One explanation for this provided by @Maki2007a is that judgment tasks are more easily influenced by outside factors such as the availability heuristic. Thus, it may be that the act of viewing the cue-target pair together at the time of judgment interferes with individuals' ability to consider other target words that are related to the cue, thereby inflating (or reducing) the perceived relatedness between the items [@Maki2007a]. Indeed, work by [@Nelson1991] has shown this to be the case when eliciting judgments of learning, as JOLs made after a delay tend to be more accurate relative to those made immediately in the presence of the studied information. Further, the influence of indirect relations and their potential interaction on judgments have not been investigated.

The present study expanded upon previous JAM studies by examining recall and judgments for three types of judgments of relatedness (JORs) with the goal of exploring the underlying memory network that is used for each of these cognitive processes as described above. To date, no study has investigated the how three types of concept information affect these judgment and recall processes within the context of one unified study. As such, we tested four hypotheses, which were based upon previous research on JAM and semantic memory models.

First, we sought to expand upon previous @Maki2007, @Maki2007a, @Buchanan2010, and @Valentine2013 research by including three types of JORs in one experiment, while replicating JAM bias and sensitivity findings. We used the SWOW norms calculating direct and indirect relations to best capture the continuum of similarity between concepts. These values were used to predict each type of JOR, and we calculated average slope and intercept values for each participant. We expected to find slope and intercept values that were significantly different from zero. Though the three types of word relations are distinct from one another, we should expect to find slopes and intercepts for semantic and thematic JORs to be within the range of previous JAM findings if these memory systems are interconnected. We also examined the frequency of each predictor being the strongest variable to predict an individual judgment condition. Thus, we are interested in exploring whether judgment findings replicate across judgment type and with new measurement variables available through SWOW (rather than each individually, as tested in previous JOL and JAM publications), which expands our knowledge on how the judgment process taps into the underlying memory network.

Next, we explored the predictions from semantic network models that the relation between associations and semantics would be bidirectional in nature (i.e., both types of knowledge interconnected in memory). Therefore, we expected to find an interaction between direct and indirect association norms when predicting JORs. We used multilevel modeling to examine the interaction of these norms in relation to participant judgments.  

We then extended these analyses to include recall as the dependent variable of interest. We tested for the interaction of database norms in predicting recall by using a multilevel logistic regression, while controlling for judgment condition and rating. We expected to find that database norms would show differences in recall based on the levels of other variables (the interaction would be significant), and that ratings would also positively predict recall (i.e., words that participants thought were more related would be remembered better). Because judgment and recall are different cognitive processes, we used this hypothesis to examine how memory networks may be differently interactive for memory in comparison to judgment. 

Finally, we examined if the judgment slopes from Hypothesis 1 would be predictive of recall. Whereas the recall model used to test our third hypothesis examined the direct relationship of word relatedness on recall, the goal of this analysis was to explore whether participant sensitivity to word relatedness was could also be used to predict recall. For this analysis, we used a multilevel logistic regression to control for multiple judgment slope conditions. This hypothesis combines both cognitive processes into one analysis, to explore how judgment ability (i.e., slopes) would impact the memory process.

# Method

## Participants

```{r data_import, include = F}

pilot = read.csv("data/Melted Data pilot.csv", stringsAsFactors = F)
thesis = read.csv("data/Melted Data Thesis.csv", stringsAsFactors = F)

thesis$Partno = thesis$Partno + 1000
pilot$where = "pilot"
thesis$where = "thesis"

colnames(pilot)[2] = "Judgment"

master = rbind(pilot, thesis)
```

A power analysis was conducted using the *simR* package in *R* [@Green2016]. This package uses simulations to generate power estimates for mixed linear models created from the *lme4* package in *R* [@Bates2015]. The results of this analyses suggested a minimum of 35 participants would be required to detect an effect. However, because power often tends to be underestimated, we extended participant recruitment as funding permitted. The data in this experiment was collected in two waves recruiting from Amazon's Mechanical Turk, which is a website that allows individuals to host projects and connects them with a large pool of respondents who complete them for small amounts of money [@Buhrmester2011]. In the first wave, a total of `r length(unique(master$Partno[master$where == "pilot"]))` participants were recruited, and in the second wave, `r length(unique(master$Partno[master$where == "thesis"]))` participants were recruited. Participant responses were screened for a basic understanding of the study's instructions. Responses were rejected for participants who entered related words when numerical judgment responses were required, and for participants who responded to the cue words during the recall phase with sentences or phrases instead of individual words. Those that completed the study correctly were compensated \$1.00 for their participation in wave one, and \$2.00 for their participation in wave two. The second wave of participants was sponsored by graduate thesis funding provided by the Missouri State University Graduate College. 

## Materials

The stimuli used were `r length(unique(master$Word.Pair))` words pairs of varying relatedness which were created from the @Buchanan2013 word norm database and website. These pairs were evenly split into sixty-three for wave one and wave two of the study.  Pairs were originally selected by using forward strength [FSG; @Nelson2004], semantic feature overlap cosine values (COS) from @Buchanan2013, and Latent Semantic Analysis cosine values [@Landauer1998; @Landauer1997] based on previous research on how word pair psycholinguistic variables overlap [@Maki2008]. The selected stimuli included a range of values for each variable. Table \@ref(tab:stim-table) displays stimuli averages, *SD*, and ranges. A complete list of stimuli can be found at http://osf.io/y8h7v. 

The stimuli were arranged into three blocks for each judgment condition described below wherein each block contained 21 word pairs. Due to limitations of the available stimuli, blocks were structured so that each one contained seven word pairs of low (0-.33), medium (.34-.66), and high (.67-1.00) COS relatedness. Pairs with low, medium, and high FSG and LSA were then selected, when avaliable. Given the measurement questions raised in the introduction, the direct association from the SWOW norms will be used as the measure of first order association. Given @DeDeyne2013a's work on continuous association, the response set from all three responses were used. The direct association provided in these norms is calculated as the number of participants who provided the target to the cue divided by the number of possible answers (i.e., participants * responses). This calculation, therefore, has an upper limit of approximately ~33%, even if every participant listed a target word to a cue. The JOR task assumes the range of direct association is 0 to 100 (or 0-1 proportion), and the SWOW direct association (DA) was normalized using: 

$$\frac{DA - Min(DA)} {Max(DA) - Min(DA)}$$

Indirect association (IA) was calculated by comparing the distribution of responses for each concept. Therefore, if the concepts were *bird* and *feather*, the two association sets were combined and the cosine between the response frequencies was calculated. Cosine indicates a measure of overlap in the response distributions, where 0 indicates no overlapping responses, while 1 indicates perfectly overlapping response frequencies [see @Buchanan2019 for more on cosine feature overlap]. DA and IA averages are provided in \@ref(tab:stim-table). The study was built online using Qualtrics, and three surveys were created to counter-balance the order in which judgment conditions appeared. Each word pair appeared counter-balanced across each judgment condition, and stimuli were randomized within each block.

```{r direct_association, include = F}

swow = read.csv("data/strength.SWOW-EN.R123.csv", stringsAsFactors = F)

swow$pair = paste(swow$cue, swow$response, sep = "_")
swow$pair = tolower(swow$pair)
swow$cue = tolower(swow$cue)
swow$response = tolower(swow$response)
swow = swow[-c(1001283) , ] #remove the dup

master$pair = gsub("...", "_", master$Word.Pair, fixed = T)
master$pair = tolower(master$pair)

master2 = merge(master, swow, by = "pair", all.x = T)
nrow(master2)
##red yellow is duplicating - removing the duplicate line from swow

#normalize SWOW as the score is not number of participants who said X / number of participants, but instead is number of participants who said X / number of answers
#normalized = (x-min(x)) / (max(x) - min(x))

master2$swow_fsg = (master2$R123.Strength - min(master2$R123.Strength, na.rm = T)) / (max(master2$R123.Strength, na.rm = T) - min(master2$R123.Strength, na.rm = T))

master2$swow_fsg[is.na(master2$swow_fsg)] = 0
```

```{r indirect_association, include = F}

paired = unique(master2$pair)
paired = as.data.frame(paired)
paired$paired = as.character(paired$paired)
paired$cosine = NA

for (i in 1:length(paired$paired)){ #loop over each word
  
  words = strsplit(paired$paired[i], "_")
  temp1 = swow[tolower(swow$cue) == words[[1]][1], ]
  temp2 = swow[tolower(swow$cue) == words[[1]][2], ]
  
  temp_merge = merge(temp1, temp2, by = "response", all = T)
  temp_merge$R123.Strength.x[is.na(temp_merge$R123.Strength.x)] = 0
  temp_merge$R123.Strength.y[is.na(temp_merge$R123.Strength.y)] = 0
  
  paired$cosine[i] = lsa::cosine(temp_merge$R123.Strength.x, temp_merge$R123.Strength.y)

}

#merge with  our data
colnames(paired)[1] = "pair"
master3 = merge(paired, master2, by = "pair")
```

```{r stim-table, echo=FALSE, results='asis'}
##create blank table
tableprint = matrix(NA, nrow = 5, ncol = 4)
colnames(tableprint) = c("Variable", "Low", "Average", "High")

##create a unique dataset
stimuli = master3[!duplicated(master3$pair), ]

##add cosine labels
stimuli$range = NA
stimuli$range[stimuli$COS < .33] = "low"
stimuli$range[stimuli$COS >= .33 & stimuli$COS < .66] = "med"
stimuli$range[stimuli$COS >= .66] = "high"
stimuli$range = factor(stimuli$range,
                       levels = c("low", "med", "high"))

##calculate means and sds
COSM = tapply(stimuli$COS, stimuli$range, mean)
COSSD = tapply(stimuli$COS, stimuli$range, sd)
tableprint[ 1, ] = c("Semantic Feature Overlap COS", 
                     pasteit(COSM[1], COSSD[1]),
                     pasteit(COSM[2], COSSD[2]),
                     pasteit(COSM[3], COSSD[3]))

FSGM = tapply(stimuli$FSG, stimuli$range, mean)
FSGSD = tapply(stimuli$FSG, stimuli$range, sd)
tableprint[ 2, ] = c("Forward Strength FSG", 
                     pasteit(FSGM[1], FSGSD[1]),
                     pasteit(FSGM[2], FSGSD[2]),
                     pasteit(FSGM[3], FSGSD[3]))

LSAM = tapply(stimuli$LSA, stimuli$range, mean)
LSASD = tapply(stimuli$LSA, stimuli$range, sd)
tableprint[ 3, ] = c("Latent Semantic Analysis LSA", 
                     pasteit(LSAM[1], LSASD[1]),
                     pasteit(LSAM[2], LSASD[2]),
                     pasteit(LSAM[3], LSASD[3]))                     
                     
DAM = tapply(stimuli$swow_fsg, stimuli$range, mean)
DASD = tapply(stimuli$swow_fsg, stimuli$range, sd)
tableprint[ 4, ] = c("Direct Association DA", 
                     pasteit(DAM[1], DASD[1]),
                     pasteit(DAM[2], DASD[2]),
                     pasteit(DAM[3], DASD[3]))                     
                     
IAM = tapply(stimuli$cosine, stimuli$range, mean)
IASD = tapply(stimuli$cosine, stimuli$range, sd)
tableprint[ 5, ] = c("Direct Association IA", 
                     pasteit(IAM[1], IASD[1]),
                     pasteit(IAM[2], IASD[2]),
                     pasteit(IAM[3], IASD[3]))  

kable(tableprint, "latex", booktabs = T, row.names = F, 
      caption = "Summary Statistics for Stimuli") %>%  
  add_header_above(c(" " = 1, "Semantic Feature Overlap COS" = 3)) %>% 
  add_footnote("$Note$. Standard deviation values are in parentheses.", notation="none", escape = F)
```

## Procedure

The present study was divided into three phases. In the first phase, JORs were elicited by presenting participants with word pairs and asking them to make judgments of how related they believed the words in each pair to be. This judgment phase consisted of three blocks of 21 word pairs which corresponded to one of three types of described word pair relationships: associative, semantic, or thematic. Each block was preceded by a set of instructions explaining one of the three types of relationships, and participants were provided with examples which illustrated the type of relationship to be judged. Participants were then presented with the word pairs to be judged. The associative block began by explaining associative memory and the role of free association tasks. Participants were provided with examples of both strong and weak associates. For example, *lost* and *found* and were presented as an example of a strongly associated pair, while *article* was paired with *newspaper*, *the*, and *clothing* to illustrate that words can have many weak associates. The semantic judgment block provided participants with a brief overview of how words are related by meaning and showed examples of concepts with both high and low feature overlap. *Tortoise* and *turtle* were provided as an example of two concepts with significant overlap. Other examples were then provided to illustrate concepts with little or no overlap. For the thematic judgments, participants were provided with an explanation of thematic relatedness. *Tree* is explained to be related to *leaf*, *fruit*, and *branch*, but not *computer*. In each judgment, participants were then given three concepts (*lost*, *old*, *article*) and were asked to come up with words that they felt were related to that type of relation. 

After viewing the examples at the start of the block, participants completed the JOR task. Each block contained a set of instructions which were contingent upon the type of JOR being elicited. For example, instructions in the associative block asked participants to estimate how many individuals out of 100 they expect would respond to the cue word with a given target, instructions for semantic JORs asked participants to indicate the percent of features shared between two concepts, and instructions for the thematic JOR task asked participants to base ratings on how likely to words would be used together in the same story. The complete experiment can be found at http://osf.io/y8h7v, which contains the exact instructions given to participants for each block and displays the structure of the study. All instructions were modeled after @Buchanan2010 and @Valentine2013.

In accordance with previous work on JOLs and JAM, participants made JOR ratings using a scale of zero to one hundred, with zero indicating no relationship, and one hundred indicating a perfect relationship. Participants typed their responses into the survey. Once completed, participants then completed the remaining judgment blocks in the same manner. Each subsequent judgment block changed the type of JOR being made. Three versions of the study were created, which counter-balanced the order in which the judgment blocks appeared, and participants were randomly assigned to a survey version. This resulted in each word pair receiving a relatedness judgments on each of the three types relationships. 

After completing the judgment phase, participants were then presented with a short distractor task to account for recency effects. In this section, participants were presented with a randomized list of the fifty U.S. states and were asked to arrange them in alphabetical order. This task was timed to last two minutes. Once time had elapsed, participants automatically progressed to the final phase, which consisted of a cued-recall task. Participants were presented with each of the sixty-three cue words from the judgment phase and were asked to complete each word pair by responding with the correct target word. Participants were informed that they would not be penalized for guessing. The cued-recall task included all stimuli in a random order.

# Results

```{r data-set-creation, eval=FALSE, include=FALSE}
##check out our 2 datasets folder to see how these datasets were made
##we moved the data into long format, and combined them all together by judgment and recall
##then we combined these data with psycholinguistic norms (in excel, if I remember correctly)
##then we merged them together in this file 

#scaling judgments
master3$Judged.Value = master3$Judged.Value / 100

#clean up judgment type
master3$Judgment = gsub("semantic", "Semantic", master3$Judgment)
master3$Judgment = gsub("associative", "Associative", master3$Judgment)
master3$Judgment = gsub("thematic", "Thematic", master3$Judgment)
```

```{r data-screening, include=FALSE}
##accuracy
summary(master3)
##fixing judged value
badjudge = sum(!is.na(master3$Judged.Value[ master3$Judged.Value > 1 ]))
master3$Judged.Value[ master3$Judged.Value > 1 ] = NA
summary(master3$Judged.Value)

##missing data will be excluded below
missingness = table("judge" = is.na(master3$Judged.Value), "recall" = is.na(master3$Recall))

##outliers based on recall and judgment
mahal = mahalanobis(master[ , c(6,7)],
                    colMeans(master[ , c(6,7)], na.rm = TRUE),
                    cov(master[ , c(6,7)], use = "pairwise.complete.obs"))
cutoff = qchisq(1-.001, ncol(master[ , c(6,7)]))
cutoff;ncol(master[ , c(6,7)])
summary(mahal < cutoff)
noout = subset(master3, mahal < cutoff)

##additivity
cor(noout[ , c(2,6,7,55)], use = "pairwise.complete.obs")

##descriptive stats
meanJno = tapply(noout$Judged.Value, noout$Judgment, mean, na.rm = T)
meanRno = tapply(noout$Recall, noout$Judgment, mean, na.rm = T)
sdJno = tapply(noout$Judged.Value, noout$Judgment, sd, na.rm = T)
sdRno = tapply(noout$Recall, noout$Judgment, sd, na.rm = T)
```

## Data Processing and Descriptive Statistics

First, the results from the recall phase of the study were coded as zero for incorrect responses, one for correct responses, and NA for participants who did not complete the recall section (all or nearly all responses were blank). All word responses to judgment items were deleted and set to missing data. The final dataset was created by splitting the initial data file into six sections (one for each of the three experimental blocks and their corresponding recall scores). Each section was individually melted using the *reshape* package in *R* [@Wickham2007] and was written as a csv file. The six output files were then combined to form the final dataset. Code is available on our OSF page embedded inline with the manuscript in an *R* markdown document written with the *papaja* package [@Aust2017]. With `r length(unique(noout$Partno))` participants, the dataset in long format included `r length(unique(noout$Partno)) * 63` rows of potential data (i.e., `r length(unique(noout$Partno))` participants \* 63 JORs). `r badjudge` out of range JOR data points (> 100) were corrected to NA. Missing data for JORs or recall were then excluded from the analysis, which included word responses to judgment items (i.e., responding with *cat* instead of a number when prompted to provide a JOR). These items usually excluded a participant from receiving Amazon Mechanical Turk payment, but were included in the datasets found online. In total,  data points`r sum(missingness[c(2,3,4)])` were excluded (`r missingness[2]` JOR only, `r missingness[3]` recall only, `r missingness[4]` both), leading to a final *N* of `r missingness[1]` observations. Recall and JOR values were then screened for outliers using Mahalanobis distance at *p* < .001, and no outliers were found [@Tabachnick2012]. To screen for multicollinearity, we examined correlations between judgment items, DA, and IA. All correlations were *r*s < .26.

```{r ANOVAs, echo=FALSE}
####set up for ANOVA####
library(reshape)

##load in master data
ANOVA.dat = read.csv("Melted Data.csv")
colnames(ANOVA.dat)[2] = "Judgment.task"
colnames(ANOVA.dat)[4] = "Judgment"

##fix out of range JORs
ANOVA.dat$Judgment[ ANOVA.dat$Judgment > 100 ] = NA

##remove outliers
mahal = mahalanobis(ANOVA.dat[ , c(4,5)],
                    colMeans(ANOVA.dat[ , c(4,5)], na.rm = TRUE),
                    cov(ANOVA.dat[ , c(4,5)], use = "pairwise.complete.obs"))
cutoff = qchisq(1-.001, ncol(ANOVA.dat[ , c(4,5)]))
cutoff;ncol(ANOVA.dat[ , c(4,5)])
summary(mahal < cutoff)
dat.noout = subset(ANOVA.dat, mahal < cutoff)

dat.noout$Judgment = dat.noout$Judgment/100

##capitalize things
dat.noout$Judgment.task = as.numeric(dat.noout$Judgment.task)

dat.noout$Judgment.task = factor(dat.noout$Judgment.task,
                           levels = c(1, 2, 3),
                           labels = c("Associative", "Semantic", "Thematic"))

##melt the data
long.dat = melt(dat.noout, id = c("Partno", "Judgment.task", "Word.Pair", "COS", "LSA", "FSG"))

##rename columns
colnames(long.dat)[2] = "Judgment.Type" 
colnames(long.dat)[7] = "Task"
colnames(long.dat)[8] = "Score"

long.dat$Score = long.dat$Score * 100 ##put on correct scale

##make Judgment and recall subsets for ANOVAs
judge = subset(long.dat,
               long.dat$Task == "Judgment")

recall = subset(long.dat,
                long.dat$Task == "Recall")

##tried running an ANOVA with the screened data and it blows up
##not all subjects have observations for each pair type
##to fix that, I'm dropping subjects that don't have responses for all 63 pairs.

temp = as.data.frame(table(judge$Partno))
temp2 = as.data.frame(table(recall$Partno))

#Returns only the subjects that have less than number of levels (Missing data)
nrow(temp[temp$Freq < 63, ]) ##32 subjects are missing judgment observations
nrow(temp2[temp2$Freq < 63, ]) ##the same 32 are missing observations for recall.

##lets drop these people and see what happens
keep_peoples = subset(temp,
                      temp$Freq == 63)

keep_peoples2 = keep_peoples$Var1 ##returns a factor
keep_peoples2 = as.numeric(as.character(keep_peoples2)) ##convert it to a number

##final judgement data
judge2 = judge[judge$Partno %in% keep_peoples2, ] ##drop everyone with less than 63 observations

table(judge2$Judgment.Type)

##final recall data
recall2 = recall[recall$Partno %in% keep_peoples2, ] ##drop those same people for the recall data

table(recall2$Judgment.Type)

####ANOVAs####
##judgments
judge.model1 = ezANOVA(data = judge2,
        wid = Partno,
        within = Judgment.Type,
        dv = Score,
        type = 3)

judge.model1 ##sig difference between judgment ratings across groups

source("pairwise_t.R")

judge_posthoc = pairwise.t.test.with.t.and.df(judge2$Score, judge2$Judgment.Type, p.adjust.method = "bonferroni", paired = F)

judge_posthoc

##get t value
judge_posthoc[[5]]

##recall
recall.model1 = ezANOVA(data = recall2,
        wid = Partno,
        within = Judgment.Type,
        dv = Score,
        type = 3)

recall.model1 ##no sig difference for recall

##anova statistics
##JORS
jf = judge.model1$ANOVA$F ##get f statistic

j.eta = judge.model1$ANOVA$ges ##partial eta squared

##Recall
rf = recall.model1$ANOVA$F

r.eta = recall.model1$ANOVA$ges
```

The mean JOR for the associative condition (*M* = `r apa(meanJno[1], 2)`, *SD* = `r apa(sdJno[1], 2)`) was lower than the semantic (*M* = `r apa(meanJno[2], 2)`, *SD* = `r apa(sdJno[2], 2)`) and thematic (*M* = `r apa(meanJno[3], 2)`, *SD* = `r apa(sdJno[3], 2)`) conditions. Recall averaged over 60% for all three conditions: associative *M* = `r apa(meanRno[1]*100, 2)`, *SD* = `r apa(sdRno[1]*100, 2)`; semantic *M* = `r apa(meanRno[2]*100, 2)`, *SD* = `r apa(sdRno[2]*100, 2)`; thematic *M* = `r apa(meanRno[3]*100, 2)`, *SD* = `r apa(sdRno[3]*100, 2)`. To confirm these patterns, two one-way ANOVAs were used to test for differences between mean JOR rates and mean correct percent recall for each of the three judgment conditions. Overall, a significant main effect of judgment type was detected for JORs (*F* (2, 144) = `r apa(jf, 2)`), *p* < .001, $\eta$*p^2^* = `r apa(j.eta, 2)`), and a post-hoc pairwise *t*-test indicated that all comparisons were signifcant (*t*s > 3.48, *p*s < .002) No significant difference was found between groups for mean correct recall (*F* (2, 144) = `r apa(rf, 2)`), *p* = .112, $\eta$*p^2^* = `r apa(r.eta, 2)`).

## JAM Slope Bias and Sensitivity

```{r hyp1-table1, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
##are average slope and intercept values greater than zero?
##are they within the range of previous work, although maybe controlling for other variables might change them a bit
##frequency of strongest predictor 

####try a loop####
##setup
persontable = matrix(NA,
                     nrow=length(names(table(noout$Partno))),
                     ncol=5+4+4)
colnames(persontable) = c("Partno", "AIntercept", "ACOS", "ALSA", "AFSG",
                          "SIntercept", "SCOS", "SLSA", "SFSG",
                          "TIntercept", "TCOS", "TLSA", "TFSG")

simnum = 1

##create the right scaling 
noout$Judged.Value2 = noout$Judged.Value/100

for ( person in names(table(noout$Partno)) ){ ##loop over participants
  
  temp1 = subset(noout, Partno == person & Judgment == "associative")
  temp2 = subset(noout, Partno == person & Judgment == "semantic")
  temp3 = subset(noout, Partno == person & Judgment == "thematic")
  
  persontable[ simnum , 1] = person
  
  if (nrow(temp1) > 9) {
    model9 = lm(Judged.Value2 ~ COS + LSA + FSG, data = temp1)
    persontable[ simnum , 2:5] = model9$coefficients
  }
  
  if(nrow(temp2) > 9) {
    model10 = lm(Judged.Value2 ~ COS + LSA + FSG, data = temp2)
    persontable[ simnum , 6:9] = model10$coefficients
  }
  
  if(nrow(temp3) > 9) {
    model11 = lm(Judged.Value2 ~ COS + LSA + FSG, data = temp3)
    persontable[ simnum , 10:13] = model11$coefficients
  }
  
  simnum = simnum + 1
  
}

##set up the output
people = apply(persontable, 2, as.numeric)
people = as.data.frame(people) 

####hyp 1 are these greater than zero####
##single sample t-test for each column, focus on the effect size

hyp1results = apply(people[ , -1], 2, function (x) { d.single.t(m = mean(x, na.rm = T), 
                                u = 0,
                                sd = sd(x, na.rm = T),
                                n = sum(!is.na(x)),
                                a = .05)})

##what happened? make a table
##create a blank table
tableprint = matrix(NA, nrow = 12, ncol = 8)
colnames(tableprint) = c("Variable", "$M$", "$SD$", "$t$", "$df$", "$p$", "$d$", "$95 CI$")

tableprint[1 , ] = c("Associative Intercept", apa(hyp1results$AIntercept$m, 3, F),
                     apa(hyp1results$AIntercept$sd, 3, F), apa(hyp1results$AIntercept$t, 3, T),
                     hyp1results$AIntercept$df, p.value(hyp1results$AIntercept$p),
                     apa(hyp1results$AIntercept$d, 3, T), paste(apa(hyp1results$AIntercept$dlow, 3, T), 
                                                     "-", apa(hyp1results$AIntercept$dhigh, 3, T)))

tableprint[2 , ] = c("Associative COS", apa(hyp1results$ACOS$m, 3, F),
                     apa(hyp1results$ACOS$sd, 3, F), apa(hyp1results$ACOS$t, 3, T),
                     hyp1results$ACOS$df, p.value(hyp1results$ACOS$p),
                     apa(hyp1results$ACOS$d, 3, T), paste(apa(hyp1results$ACOS$dlow, 3, T), 
                                                     "-", apa(hyp1results$ACOS$dhigh, 3, T)))

tableprint[3 , ] = c("Associative FSG", apa(hyp1results$AFSG$m, 3, F),
                     apa(hyp1results$AFSG$sd, 3, F), apa(hyp1results$AFSG$t, 3, T),
                     hyp1results$AFSG$df, p.value(hyp1results$AFSG$p),
                     apa(hyp1results$AFSG$d, 3, T), paste(apa(hyp1results$AFSG$dlow, 3, T), 
                                                     "-", apa(hyp1results$AFSG$dhigh, 3, T)))

tableprint[4 , ] = c("Associative LSA", apa(hyp1results$ALSA$m, 3, F),
                     apa(hyp1results$ALSA$sd, 3, F), apa(hyp1results$ALSA$t, 3, T),
                     hyp1results$ALSA$df, p.value(hyp1results$ALSA$p),
                     apa(hyp1results$ALSA$d, 3, T), paste(apa(hyp1results$ALSA$dlow, 3, T), 
                                                     "-", apa(hyp1results$ALSA$dhigh, 3, T)))

tableprint[5 , ] = c("Semantic Intercept", apa(hyp1results$SIntercept$m, 3, F),
                     apa(hyp1results$SIntercept$sd, 3, F), apa(hyp1results$SIntercept$t, 3, T),
                     hyp1results$SIntercept$df, p.value(hyp1results$SIntercept$p),
                     apa(hyp1results$SIntercept$d, 3, T), paste(apa(hyp1results$SIntercept$dlow, 3, T), 
                                                     "-", apa(hyp1results$SIntercept$dhigh, 3, T)))

tableprint[6 , ] = c("Semantic COS", apa(hyp1results$SCOS$m, 3, F),
                     apa(hyp1results$SCOS$sd, 3, F), apa(hyp1results$SCOS$t, 3, T),
                     hyp1results$SCOS$df, p.value(hyp1results$SCOS$p),
                     apa(hyp1results$SCOS$d, 3, T), paste(apa(hyp1results$SCOS$dlow, 3, T), 
                                                     "-", apa(hyp1results$SCOS$dhigh, 3, T)))

tableprint[7 , ] = c("Semantic FSG", apa(hyp1results$SFSG$m, 3, F),
                     apa(hyp1results$SFSG$sd, 3, F), apa(hyp1results$SFSG$t, 3, T),
                     hyp1results$SFSG$df, p.value(hyp1results$SFSG$p),
                     apa(hyp1results$SFSG$d, 3, T), paste(apa(hyp1results$SFSG$dlow, 3, T), 
                                                     "-", apa(hyp1results$SFSG$dhigh, 3, T)))

tableprint[8 , ] = c("Semantic LSA", apa(hyp1results$SLSA$m, 3, F),
                     apa(hyp1results$SLSA$sd, 3, F), apa(hyp1results$SLSA$t, 3, T),
                     hyp1results$SLSA$df, p.value(hyp1results$SLSA$p),
                     apa(hyp1results$SLSA$d, 3, T), paste(apa(hyp1results$SLSA$dlow, 3, T), 
                                                     "-", apa(hyp1results$SLSA$dhigh, 3, T)))

tableprint[9 , ] = c("Thematic Intercept", apa(hyp1results$TIntercept$m, 3, F),
                     apa(hyp1results$TIntercept$sd, 3, F), apa(hyp1results$TIntercept$t, 3, T),
                     hyp1results$TIntercept$df, p.value(hyp1results$TIntercept$p),
                     apa(hyp1results$TIntercept$d, 3, T), paste(apa(hyp1results$TIntercept$dlow, 3, T), 
                                                     "-", apa(hyp1results$TIntercept$dhigh, 3, T)))

tableprint[10 , ] = c("Thematic COS", apa(hyp1results$TCOS$m, 3, F),
                     apa(hyp1results$TCOS$sd, 3, F), apa(hyp1results$TCOS$t, 3, T),
                     hyp1results$TCOS$df, p.value(hyp1results$TCOS$p),
                     apa(hyp1results$TCOS$d, 3, T), paste(apa(hyp1results$TCOS$dlow, 3, T), 
                                                     "-", apa(hyp1results$TCOS$dhigh, 3, T)))

tableprint[11 , ] = c("Thematic FSG", apa(hyp1results$TFSG$m, 3, F),
                     apa(hyp1results$TFSG$sd, 3, F), apa(hyp1results$TFSG$t, 3, T),
                     hyp1results$TFSG$df, p.value(hyp1results$TFSG$p),
                     apa(hyp1results$TFSG$d, 3, T), paste(apa(hyp1results$TFSG$dlow, 3, T), 
                                                     "-", apa(hyp1results$TFSG$dhigh, 3, T)))

tableprint[12 , ] = c("Thematic LSA", apa(hyp1results$TLSA$m, 3, F),
                     apa(hyp1results$TLSA$sd, 3, F), apa(hyp1results$TLSA$t, 3, T),
                     hyp1results$TLSA$df, p.value(hyp1results$TLSA$p),
                     apa(hyp1results$TLSA$d, 3, T), paste(apa(hyp1results$TLSA$dlow, 3, T), 
                                                     "-", apa(hyp1results$TLSA$dhigh, 3, T)))

##of course I figured out an easier way later... 

apa_table(as.data.frame(tableprint), 
          align = c("l", rep("c", 7)), 
          caption = "Summary Statistics for Hypothesis 1 t-Tests",
          note = "Confidence interval for $d$ was calculated using the non-central $t$-distribution. ",
          col.names = c("Variable", "$M$", "$SD$", "$t$", "$df$", "$p$", "$d$", "$95\\% CI$"),
          escape = FALSE
 )

```

First, we sought to replicate bias and sensitivity findings from previous research while expanding the JAM function to include judgments based on three types of memory. FSG, COS, and LSA were used to predict each type of relatedness judgment. JOR values were divided by 100, so as to place them on the same scale as the database norms. Slopes and intercepts were then calculated for each participant's ratings for each of the three JOR conditions, as long as they contained at least nine data points out of the twenty-one that were possible. Single sample *t*-tests were then conducted to test if slope and intercept values significantly differed from zero. See Table \@ref(tab:hyp1-table1) for means and standard deviations. Slopes were then compared to the JAM function, which is characterized by high intercepts (between 40 and 60 on a 100 point scale) and shallow slopes (between 20 and 40). Because of the scaling of our data, to replicate this function, we should expect to find intercepts ranging from .40 to .60 and slopes in the range of 0.20. to 0.40. Intercepts for associative, semantic, and thematic JORs were each significant, and all fell within or near the expected range. Overall, thematic JORs had the highest intercept at `r apa(hyp1results$TIntercept$m, 3, F)`, while JORs elicited in the associative condition had the lowest intercept at `r apa(hyp1results$AIntercept$m, 3, F)`. 

The JAM slope was successfully replicated for FSG in the associative JOR condition, with FSG significantly predicting association, although the slope was slightly higher than expected at `r apa(hyp1results$AFSG$m, 3, F)`. COS and LSA did not significantly predict association. For semantic judgments, each of the three database norms were significant predictors. However, JAM slopes were not replicated for this judgment type, as FSG had the highest slope at `r apa(hyp1results$SFSG$m, 3, F)`, followed by LSA `r apa(hyp1results$SLSA$m, 3, F)`, and then COS `r apa(hyp1results$SCOS$m, 3, F)`. These findings were mirrored for thematic JORs, as each database norm was a significant predictor, yet slopes for each predictor fell below range of the expected JAM slopes. Again, FSG had the highest slope, this time just out of range at `r apa(hyp1results$TFSG$m, 3, F)`, followed closely by LSA at `r apa(hyp1results$TLSA$m, 3, F)`. Interestingly, COS slopes were found to be negative for this judgment condition, `r apa(hyp1results$TCOS$m, 3, F)`. Overall, although JAM slopes were not successfully replicated in each JOR condition, the high intercepts and shallow slopes present across conditions are still indicative of overconfidence and insensitivity in participant JORs. 

```{r hyp1-part2, include=FALSE}
##group them all by type
##dropping partno column
people_thematic = people[ , -c(1:9)]
people_semantic = people[ , -c(1:5, 10:13)]
people_associative = people[ , -c(1, 6:13)]

##getting absolute values
people_thematic2 = abs(people_thematic)
people_semantic2 = abs(people_semantic)
people_associative2 = abs(people_associative)

##getting rid of the intercepts
people_thematic3 = people_thematic2[ , -1]
people_semantic3 = people_semantic2[ , -1]
people_associative3 = people_associative2[ , -1]

##finding max coefficents for each row
thematic_output = names(people_thematic3)[max.col(people_thematic3, ties.method="first")]
thematic_output = as.data.frame(thematic_output)

semantic_output = names(people_semantic3)[max.col(people_semantic3, ties.method="first")]
semantic_output = as.data.frame(semantic_output)

associative_output = names(people_associative3)[max.col(people_associative3, ties.method="first")]
associative_output = as.data.frame(associative_output)

##sticking it all together
combined = cbind(people$Partno, associative_output, semantic_output, thematic_output)
colnames(combined) = c("Partno", "Associative", "Semantic", "Thematic")

##rename the things
combined$Associative = factor(combined$Associative,
                           labels = c("COS", "FSG", "LSA"))
combined$Semantic = factor(combined$Semantic,
                           labels = c("COS", "FSG", "LSA"))
combined$Thematic = factor(combined$Thematic,
                           labels = c("COS", "FSG", "LSA"))
summary(combined[ , -1])

##what are the percentages on these?
library(memisc)
p1 = percent(combined$Associative)
p1
p2 = percent(combined$Semantic)
p2
p3 = percent(combined$Thematic)
p3

```

Additionally, we examined the frequency that each predictor variable was the strongest predictor for each of the three JOR conditions. For the associative condition, FSG was the strongest predictor for `r apa(p1[2],1)`% of the participants, with COS and LSA being the strongest for only `r apa(p1[1],1)`% and `r apa(p1[3],1)`% of participants respectively. These differences were less distinct when examining the semantic and thematic JOR conditions. In the semantic condition, FSG was highest at `r apa(p2[2], 1)`% of participants, LSA was second at `r apa(p2[3],1)`%, and COS was least likely at `r apa(p2[1],1)`%. Finally, in the thematic condition, LSA was most likely to be the strongest predictor with `r apa(p3[3],1)`% of participants, with FSG being the second most likely at `r apa(p3[2],1)`%, and COS again being least likely at `r apa(p3[1],1)`%. Interestingly, in all three conditions, COS was least likely to be the strongest predictor, even in the semantic condition. Therefore, these results provide evidence of the nature of judgments on the memory network as each judgment type appeared to tap each tier differently, suggesting a three-part system, rather than one large, encompassing memory network. 

## Interaction between Norms when Predicting Judgments of Relatedness
```{r hyp2-table, echo=FALSE, results='asis'}
##interaction in predicting judgments, use MLM to control for random participant intercepts 

####mean center the predictor variables####
noout$ZCOS = scale(noout$COS, scale = F)
noout$ZLSA = scale(noout$LSA, scale = F)
noout$ZFSG = scale(noout$FSG, scale = F)

##create the right scaling 
noout$Judged.Value2 = noout$Judged.Value/100

####overall model####
library(nlme)
library(MuMIn)
overallh2 = lme(Judged.Value2 ~ Judgment + 
               ZCOS * ZLSA * ZFSG, 
             data = noout, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|Partno)
#summary(overallh2) ##everything is significant
#r.squaredGLMM(overallh2) ##use R2m to talk about variance predicted


####power test####
#library(nlmeU)
#library(simr)
#Pwr(overallh2)

#powerest = rep(NA, 99)
#samplesizetest = seq(1*63, 100*63, 62) ##set up sample sizes to test
#sim = 1
#for (i in samplesizetest){

#smalldata = sample(noout, i)

#testh2 = lme(Judged.Value2 ~ Judgment + 
#               ZCOS * ZLSA * ZFSG, 
#             data = smalldata, 
#             method = "ML", 
#             na.action = "na.omit",
#             random = ~1|Partno)
#powerest[sim] = Pwr(testh2)[9,5]
#sim = sim + 1
#}

#powerest
#{plot(powerest)
#abline(h = .80)}


####table set up####
##create a blank table
tableprint = matrix(NA, nrow = 22, ncol = 5)
colnames(tableprint) = c("Variable", "$beta$", "$SE$", "$t$", "$p$")

tableprint[1:10 , 2:5] = summary(overallh2)$tTable[ , -3]
tableprint[1:10, 1] = c("Intercept", "Semantic Judgments", "Thematic Judgments", dimnames(summary(overallh2)$tTable[ , -3])[[1]][4:10] )

####examine three way interaction by breaking down the COS variable first####
##setup
noout$ZCOS_low = noout$ZCOS + sd(noout$ZCOS, na.rm = TRUE)
noout$ZCOS_high = noout$ZCOS - sd(noout$ZCOS, na.rm = TRUE)

##low cosine
lowcos = lme(Judged.Value2 ~ Judgment +
               ZCOS_low * ZLSA * ZFSG, 
             data = noout, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|Partno)
#summary(lowcos) ##two way interaction is significant
tableprint[11:13 , 2:5] = summary(lowcos)$tTable[ c(5,6,9), -3]
tableprint[11:13, 1] = c("Low COS ZLSA", 
                         "Low COS ZFSG", 
                         "Low COS ZLSA:ZFSG")

##high cosine
highcos = lme(Judged.Value2 ~ Judgment +
                ZCOS_high * ZLSA * ZFSG, 
              data = noout, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|Partno)
#summary(highcos) ##two way interaction is significant

tableprint[14:16 , 2:5] = summary(highcos)$tTable[ c(5,6,9), -3]
tableprint[14:16, 1] = c("High COS ZLSA", 
                         "High COS ZFSG", 
                         "High COS ZLSA:ZFSG")


####examine the two way interactions by splitting LSA####
##setup
noout$ZLSA_low = noout$ZLSA + sd(noout$ZLSA, na.rm = TRUE)
noout$ZLSA_high = noout$ZLSA - sd(noout$ZLSA, na.rm = TRUE)

##low cosine, low lsa
lowcoslowlsa = lme(Judged.Value2 ~ Judgment +
               ZCOS_low * ZLSA_low * ZFSG, 
               data = noout, 
               method = "ML", 
             na.action = "na.omit",
             random = ~1|Partno)
#summary(lowcoslowlsa) ##fsg is sig
tableprint[17, ] = c("Low COS Low LSA ZFSG",
                     summary(lowcoslowlsa)$tTable[6, -3])

##low cos, avg lsa is low cosine overall model

##low cos, high lsa
lowcoshighlsa = lme(Judged.Value2 ~ Judgment +
                ZCOS_low  * ZLSA_high * ZFSG, 
              data = noout, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|Partno)
#summary(lowcoshighlsa) ##fsg is not sig
tableprint[18, ] = c("Low COS High LSA ZFSG",
                     summary(lowcoshighlsa)$tTable[6, -3])

##avg cos, low lsa
acgcoslowlsa = lme(Judged.Value2 ~ Judgment +
               ZCOS * ZLSA_low * ZFSG, 
               data = noout, 
               method = "ML", 
             na.action = "na.omit",
             random = ~1|Partno)
#summary(acgcoslowlsa) ##fsg is sig
tableprint[19, ] = c("Avg COS Low LSA ZFSG",
                     summary(acgcoslowlsa)$tTable[6, -3])

##avg cos, avg lsa is overall model overallh2

##avg cos, high lsa
avgcoshighlsa = lme(Judged.Value2 ~ Judgment +
                ZCOS  * ZLSA_high * ZFSG, 
              data = noout, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|Partno)
#summary(avgcoshighlsa) ##fsg is sig
tableprint[20, ] = c("Avg COS High LSA ZFSG",
                     summary(avgcoshighlsa)$tTable[6, -3])

##high cos, low lsa
highcoslowlsa = lme(Judged.Value2 ~ Judgment +
                     ZCOS_high * ZLSA_low * ZFSG, 
                   data = noout, 
                   method = "ML", 
                   na.action = "na.omit",
                   random = ~1|Partno)
#summary(highcoslowlsa) ##fsg is not sig
tableprint[21, ] = c("High COS Low LSA ZFSG",
                     summary(highcoslowlsa)$tTable[6, -3])

##high cos, avg lsa is the overall high cosine model

##high cos, high lsa
highcoshighlsa = lme(Judged.Value2 ~ Judgment +
                      ZCOS_high  * ZLSA_high * ZFSG, 
                    data = noout, 
                    method = "ML", 
                    na.action = "na.omit",
                    random = ~1|Partno)
#summary(highcoshighlsa) ##fsg is sig
tableprint[22, ] = c("High COS High LSA ZFSG",
                     summary(highcoshighlsa)$tTable[6, -3])
tableprint[ , 2] = apa(as.numeric(tableprint[ , 2]),3,T) 
tableprint[ , 3] = apa(as.numeric(tableprint[ , 3]),3,T)
tableprint[ , 4] = apa(as.numeric(tableprint[ , 4]),3,T)
tableprint[ , 5] =  apply(as.data.frame(as.numeric(tableprint[ , 5])), 1, p.value)

apa_table(as.data.frame(tableprint),
          align = c("l", rep("c", 4)), 
          small = T,
          caption = "MLM Statistics for Hypothesis 2",
          note = "Database norms were mean centered. The table shows main effects and interactions for database norms at low, average, and high levels of COS and LSA when predicting participant judgments.",
          escape = FALSE
 )

```

```{r hyp2graph, echo=FALSE, fig.cap = "Simple slopes graph displaying the slope of FSG when predicting JORs at low, average, and high LSA split by low, average, and high COS. All variables were mean centered.", fig.height=6, fig.width=6}
##hyp 2 graphs
####setup####
dat = noout

##low cos
plot1 = ggplot(dat, aes(x = ZCOS_low, y = Judged.Value2)) +
  labs(x = "ZFSG", y = "Judgments") +
  scale_size_continuous(guide = FALSE) +
  geom_abline(aes(intercept = .607, slope = .663, linetype = "-1SD ZLSA")) +
  geom_abline(aes(intercept = .632, slope = .375, linetype = "Average ZLSA")) +
  geom_abline(aes(intercept = .657, slope = .087, linetype = "+1SD ZLSA")) +
  scale_linetype_manual(values = c("dotted", "dashed", "solid"),
                        breaks = c("-1SD ZLSA", "Average ZLSA", "+1SD ZLSA"),
                        name = "Simple Slope") +
  coord_cartesian(xlim = c(-.20, .60), ylim = c(.1, 1)) +
  geom_vline(xintercept = -.30) +
  geom_hline(yintercept = 0) +
  cleanup + 
  labs(title="Low ZCOS") 

##avg cos
plot2 = ggplot(dat, aes(x = ZCOS_low, y = Judged.Value2)) +
  labs(x = "ZFSG", y = "Judgments") +
  scale_size_continuous(guide = FALSE) +
  geom_abline(aes(intercept = .586, slope = .381, linetype = "-1SD ZLSA")) +
  geom_abline(aes(intercept = .603, slope = .271, linetype = "Average ZLSA")) +
  geom_abline(aes(intercept = .621, slope = .161, linetype = "+1SD ZLSA")) +
  scale_linetype_manual(values = c("dotted", "dashed", "solid"),
                        breaks = c("-1SD ZLSA", "Average ZLSA", "+1SD ZLSA"),
                        name = "Simple Slope") +
  coord_cartesian(xlim = c(-.20, .60), ylim = c(.1, 1)) +
  geom_vline(xintercept = -.30) +
  geom_hline(yintercept = 0) +
  cleanup + 
  labs(title="Average ZCOS") 

##high cos
plot3 = ggplot(dat, aes(x = ZCOS_low, y = Judged.Value2)) +
  labs(x = "ZFSG", y = "Judgments") +
  scale_size_continuous(guide = FALSE) +
  geom_abline(aes(intercept = .564, slope = .099, linetype = "-1SD ZLSA")) +
  geom_abline(aes(intercept = .575, slope = .167, linetype = "Average ZLSA")) +
  geom_abline(aes(intercept = .586, slope = .236, linetype = "+1SD ZLSA")) +
  scale_linetype_manual(values = c("dotted", "dashed", "solid"),
                        breaks = c("-1SD ZLSA", "Average ZLSA", "+1SD ZLSA"),
                        name = "Simple Slope") +
  coord_cartesian(xlim = c(-.20, .60), ylim = c(.1, 1)) +
  geom_vline(xintercept = -.30) +
  geom_hline(yintercept = 0) +
  cleanup + 
  labs(title="High ZCOS") 

# arrange plots together
legend = get_legend(plot1)
hyp2graphout <- plot_grid( plot1 + theme(legend.position="none"),
                   plot2 + theme(legend.position="none"),
                   plot3 + theme(legend.position="none"),
                   legend,
                   hjust = -1,
                   nrow = 2
)
hyp2graphout

#tiff(filename = "hyp2graph.tiff", res = 300, width = 6, 
#     height = 6, units = 'in', compression = "lzw")
#plot(hyp2graphout)
#dev.off()
```

The goal of next analysis was to test for an interaction between the three overlap norms when predicting participant JORs to examine the bidirectional network model. First, the database norms were mean centered to control for multicollinearity. The *nlme* package and *lme* function were used to calculate these analyses [@Pinheiro2017]. A maximum likelihood multilevel model was used to test the interaction between FSG, COS, and LSA when predicting JOR values, with participant number used as the random intercept factor. The type of JOR being elicited was controlled for, so as to better assess the impact of each word overlap measure regardless of JOR condition. Multilevel models were used to retain all data points (rather than averaging over items and conditions) while controlling for correlated error due to participants, which makes these models advantageous for multiway repeated measures designs [@Gelman2006]. This analysis resulted in a significant three-way interaction between FSG, COS, and LSA ($\beta$ = `r tableprint[10 , 2]`, *p* `r tableprint[10 , 5]`), which is examined below in a simple slopes analysis. Table \@ref(tab:hyp2-table) includes values for main effects, two-way, and three-way interactions.

To investigate this interaction, simple slopes were calculated for low, average, and high levels of COS. This variable was chosen for two reasons: first, it was found to be the weakest of the three predictors in hypothesis one, and second, manipulating COS would allow us to track changes across FSG and LSA. Significant two-way interactions were found between FSG and LSA at both low COS ($\beta$ = `r tableprint[13 , 2]`, *p* `r tableprint[13 , 5]`), average COS ($\beta$ =`r tableprint[9 , 2]`, *p* `r tableprint[9 , 5]`), and high COS ($\beta$ = `r tableprint[16 , 2]`, *p* = `r tableprint[16 , 5]`). A second level was then added to the analysis in which simple slopes were created for each level of LSA, allowing us to assess the effects of LSA at different levels of COS on FSG. When both COS and LSA were low, FSG significantly predicted JOR values ($\beta$ = `r tableprint[17 , 2]`, *p* `r tableprint[17 , 5]`). At low COS and average LSA, FSG decreased but still significantly predicted JORs ($\beta$ = `r tableprint[12 , 2]`, *p* `r tableprint[12 , 5]`). However, when COS was low and LSA was high, FSG was not a significant predictor ($\beta$ = `r tableprint[18 , 2]`, *p* = `r tableprint[18 , 5]`). A similar set of results was found at the average COS level. When COS was average and LSA was LOW, FSG was a significant predictor, ($\beta$ = `r tableprint[19 , 2]`, *p* `r tableprint[19 , 5]`). As LSA increased at average COS levels, FSG decreased in strength: average COS, average LSA FSG ($\beta$ = `r tableprint[16 , 2]`, *p* = `r tableprint[16 , 5]`) and average COS, high LSA FSG ($\beta$ = `r tableprint[20 , 2]`, *p* `r tableprint[20 , 5]`). This finding suggests that at low COS, LSA and FSG create a seesaw effect in which increasing levels of thematics is counterbalanced by decreasing importance of association when predicting JORs. FSG was not a significant predictor when COS was high and LSA was low (`r tableprint[21 , 2]`, *p* = `r tableprint[21, 5]`). At high COS and average LSA, FSG significantly predicted JORs ($\beta$ = `r tableprint[15 , 2]`, *p* `r tableprint[15 , 5]`), and finally when both COS and LSA were high, FSG increased and was a significant predictor of JOR values ($\beta$ = `r tableprint[22 , 2]`, *p* `r tableprint[22 , 5]`). Thus, at high levels of semantic overlap, associative and thematic overlap are complementary when predicting JOR ratings, increasing together as semantic strength increases. Figure \@ref(fig:hyp2graph) displays the three-way interaction wherein the top row of figures indicates the seesaw effect, as thematic strength increases, the predictive ability of associative overlap decreases in strength. The bottom row indicates the complementary effect where increases in LSA occur with increases in FSG predictor strength. Therefore, the cognitive process of judgment appears to be interactive in nature across these three types of memory information.

## Interaction between Norms when Predicting Recall
```{r hyp3-table, echo=FALSE, results='asis'}
##old hypothesis 1
##using recall as the DV, CV judgment type, rating, IVs three database norms
##use zscores created for hyp 2

##overall model
library(lme4)
overallh3 = glmer(Recall ~ (1|Partno) + Judgment + 
                    Judged.Value2 + ZCOS * ZLSA * ZFSG,
               data = noout,
               family = binomial,
               control = glmerControl(optimizer = "bobyqa"),
               nAGQ = 1)

#summary(overallh3)
#r.squaredGLMM(overallh3)

####table set up####
##create a blank table
tableprint = matrix(NA, nrow = 21, ncol = 5)
colnames(tableprint) = c("Variable", "$beta$", "$SE$", "$z$", "$p$")

tableprint[1:11 , 2:5] = coef(summary(overallh3))
tableprint[1:11, 1] = c("Intercept", "Semantic Judgments", "Thematic Judgments", "Judged Values", dimnames(coef(summary(overallh3)))[[1]][5:11])

#low cosine
lowcos2 = glmer(Recall ~ (1|Partno) +
                  Judgment + Judged.Value2 + ZCOS_low * ZLSA * ZFSG,
               data = noout,
               family = binomial,
               control = glmerControl(optimizer = "bobyqa"),
               nAGQ = 1)

tableprint[12:14 , 2:5] = coef(summary(lowcos2))[c(6,7,10), ]
tableprint[12:14, 1] = c("Low COS ZLSA", 
                         "Low COS ZFSG", 
                         "Low COS ZLSA:ZFSG")

##high cosine
hicos2 = glmer(Recall ~ (1|Partno) + 
                 Judgment + Judged.Value2 + ZCOS_high * ZLSA * ZFSG,
                        data = noout,
                        family = binomial,
                        control = glmerControl(optimizer = "bobyqa"),
                        nAGQ = 1)

tableprint[15:17, 2:5] = coef(summary(hicos2))[c(6,7,10), ]
tableprint[15:17, 1] = c("High COS ZLSA", 
                         "High COS ZFSG", 
                         "High COS ZLSA:ZFSG")

####splitting lsa by cosine strength####
##low cosine low lsa
lowcoslowlsa2 = glmer(Recall ~ (1|Partno) + 
                        Judgment + Judged.Value2 + ZCOS_low * ZLSA_low * ZFSG,
                      data = noout,
                      family = binomial,
                      control = glmerControl(optimizer = "bobyqa"),
                      nAGQ = 1)

tableprint[18, ] = c("Low COS Low LSA ZFSG", coef(summary(lowcoslowlsa2))[7,])

##low cosine high lsa
lowcoshighlsa2 = glmer(Recall ~ (1|Partno) + 
                        Judgment + Judged.Value2 + ZCOS_low * ZLSA_high * ZFSG,
                      data = noout,
                      family = binomial,
                      control = glmerControl(optimizer = "bobyqa"),
                      nAGQ = 1)

tableprint[19, ] = c("Low COS High LSA ZFSG", coef(summary(lowcoshighlsa2))[7,])

##high cos low lsa
highcoslowlsa2 = glmer(Recall ~ (1|Partno) + 
                        Judgment + Judged.Value2 + ZCOS_high * ZLSA_low * ZFSG,
                      data = noout,
                      family = binomial,
                      control = glmerControl(optimizer = "bobyqa"),
                      nAGQ = 1)

tableprint[20, ] = c("High COS Low LSA ZFSG", coef(summary(highcoslowlsa2))[7,])

##high cosine high lsa
highcoshighlsa2 = glmer(Recall ~ (1|Partno) + 
                        Judgment + Judged.Value2 + ZCOS_high * ZLSA_high * ZFSG,
                      data = noout,
                      family = binomial,
                      control = glmerControl(optimizer = "bobyqa"),
                      nAGQ = 1)

tableprint[21, ] = c("High COS High LSA ZFSG", coef(summary(highcoshighlsa2))[7,])

tableprint[ , 2] = apa(as.numeric(tableprint[ , 2]),3,T) 
tableprint[ , 3] = apa(as.numeric(tableprint[ , 3]),3,T)
tableprint[ , 4] = apa(as.numeric(tableprint[ , 4]),3,T)
tableprint[ , 5] = apply(as.data.frame(as.numeric(tableprint[ , 5])), 1, p.value)


apa_table(as.data.frame(tableprint),
          align = c("l", rep("c", 4)), 
          small = T,
          caption = "MLM Statistics for Hypothesis 3",
          note = "Database norms were mean centered. The table shows main effects and interactions for database norms at low, average, and high levels of COS and LSA when predicting recall.",
          escape = FALSE
 )

```

```{r hyp3graph, echo=FALSE, fig.cap = "Simple slopes graph displaying the slope of FSG when predicting recall at low, average, and high LSA split by low, average, and high COS. All variables were mean centered.", fig.height=6, fig.width=6}

##low cos
plot4 = ggplot(dat, aes(x = ZCOS_low, y = Judged.Value2)) +
  labs(x = "ZFSG", y = "Recall") +
  scale_size_continuous(guide = FALSE) +
  geom_abline(aes(intercept = 0.316, slope = 4.116, linetype = "-1SD ZLSA")) +
  geom_abline(aes(intercept = 0.136, slope = 2.601, linetype = "Average ZLSA")) +
  geom_abline(aes(intercept = -0.044, slope = 1.086, linetype = "+1SD ZLSA")) +
  scale_linetype_manual(values = c("dotted", "dashed", "solid"),
                        breaks = c("-1SD ZLSA", "Average ZLSA", "+1SD ZLSA"),
                        name = "Simple Slope") +
  coord_cartesian(xlim = c(-.20, .60), ylim = c(.1, 1)) +
  geom_vline(xintercept = -.30) +
  geom_hline(yintercept = 0) +
  cleanup + 
  labs(title="Low ZCOS") 

##avg cos
plot5 = ggplot(dat, aes(x = ZCOS_low, y = Judged.Value2)) +
  labs(x = "ZFSG", y = "Recall") +
  scale_size_continuous(guide = FALSE) +
  geom_abline(aes(intercept = 0.369, slope = 3.281, linetype = "-1SD ZLSA")) +
  geom_abline(aes(intercept = 0.301, slope = 3.084, linetype = "Average ZLSA")) +
  geom_abline(aes(intercept = 0.234, slope = 2.889, linetype = "+1SD ZLSA")) +
  scale_linetype_manual(values = c("dotted", "dashed", "solid"),
                        breaks = c("-1SD ZLSA", "Average ZLSA", "+1SD ZLSA"),
                        name = "Simple Slope") +
  coord_cartesian(xlim = c(-.20, .60), ylim = c(.1, 1)) +
  geom_vline(xintercept = -.30) +
  geom_hline(yintercept = 0) +
  cleanup + 
  labs(title="Average ZCOS") 

##high cos
plot6 = ggplot(dat, aes(x = ZCOS_low, y = Judged.Value2)) +
  labs(x = "ZFSG", y = "Recall") +
  scale_size_continuous(guide = FALSE) +
  geom_abline(aes(intercept = .421, slope = 2.44, linetype = "-1SD ZLSA")) +
  geom_abline(aes(intercept = .466, slope = 3.569, linetype = "Average ZLSA")) +
  geom_abline(aes(intercept = .511, slope = 4.692, linetype = "+1SD ZLSA")) +
  scale_linetype_manual(values = c("dotted", "dashed", "solid"),
                        breaks = c("-1SD ZLSA", "Average ZLSA", "+1SD ZLSA"),
                        name = "Simple Slope") +
  coord_cartesian(xlim = c(-.20, .60), ylim = c(.1, 1)) +
  geom_vline(xintercept = -.30) +
  geom_hline(yintercept = 0) +
  cleanup + 
  labs(title="High ZCOS") 

legend = get_legend(plot4)
hyp3graphout <- plot_grid( plot4 + theme(legend.position="none"),
                   plot5 + theme(legend.position="none"),
                   plot6 + theme(legend.position="none"),
                   legend,
                   hjust = -1,
                   nrow = 2
)
hyp3graphout

#tiff(filename = "hyp3graph.tiff", res = 300, width = 6, 
#     height = 6, units = 'in', compression = "lzw")
#plot(hyp3graphout)
#dev.off()

```

Given the results of Hypothesis 2, we then sought to extend the analysis to participant recall scores. A multilevel logistic regression was used with the *lme4* package and *glmer()* function [@Pinheiro2017], testing the interaction between FSG, COS, and LSA when predicting participant recall. As with the previous hypothesis, we controlled for JOR condition and, additionally, covaried JOR ratings. Participants were used as a random intercept factor. Judged values were a significant predictor of recall, ($\beta$ = `r tableprint[4, 2]`, *p* `r tableprint[4, 5]`) where increases in judged strength predicted increases in recall. A significant three-way interaction was detected between FSG, COS, and LSA ($\beta$ = `r tableprint[11, 2]`, *p* `r tableprint[11, 5]`). See Table \@ref(tab:hyp3-table) for main effects, two-way, and three-way interaction values. 

The same moderation process used in Hypothesis 2 was then repeated, with simple slopes first calculated at low, average, and high levels of COS. This set of analyses resulted in significant two-way interactions between LSA and FSG at low COS ($\beta$ = `r tableprint[14, 2]`, *p* `r tableprint[14, 5]`) and high COS ($\beta$ = `r tableprint[17, 2]`, *p* = `r tableprint[17, 5]`). No significant two-way interaction was found at average COS ($\beta$ = `r tableprint[10, 2]`, *p* = `r tableprint[10, 5]`). Following the design used when predicting JORs, simple slopes were then calculated for low, average, and high levels of LSA at the low and high levels of COS, allowing us to assess how FSG affects recall at varying levels of both COS and LSA. When both COS and LSA were low, FSG was a significant predictor of recall ($\beta$ = `r tableprint[18, 2]`, *p* `r tableprint[18, 5]`). At low COS and average LSA, FSG decreased from both low levels, but was still a significant predictor ($\beta$ = `r tableprint[13, 2]`, *p* `r tableprint[13, 5]`), and finally, low COS and high LSA, FSG was the weakest predictor of the three ($\beta$ = `r tableprint[19, 2]`, *p* = `r tableprint[19, 5]`). As with Hypothesis 2, LSA and FSG counterbalanced one another, wherein the increasing levels of thematics led to a decrease in the importance of association in predicting recall. At high COS and low LSA, FSG was a significant predictor ($\beta$ = `r tableprint[20, 2]`, *p* = `r tableprint[20, 5]`). When COS was high and LSA was average, FSG increased as a predictor and remained significant ($\beta$ = `r tableprint[16, 2]`, *p* `r tableprint[16, 5]`). This finding repeated when both COS and LSA were high, with FSG increasing as a predictor of recall ($\beta$ = `r tableprint[21, 2]`, *p* `r tableprint[21, 5]`). Therefore, at high levels of at high levels of semantics, thematics and association are complementary predictors of recall, increasing together and extending the findings of Hypothesis 2 to participant recall. Figure \@ref(fig:hyp3graph) displays the three-way interaction. The top left figure indicates the counterbalancing effect of recall of LSA and FSG, while the top right figure shows no differences in simple slopes for average levels of cosine. The bottom left figure indicates the complementary effects where LSA and FSG increase together as predictors of recall at high COS levels. 

## Predicting Recall with JAM Slopes

```{r hyp4-table, echo=FALSE, results='asis'}
##do the slopes from hyp 1 predict recall 

##merge recall back with slopes from the people dataset
hyp4data = merge(people, noout, by = "Partno")

##association
assoch4 = glmer(Recall ~ (1|Partno) + ACOS + ALSA + AFSG + AIntercept,
                data = hyp4data,
                family = binomial,
                control = glmerControl(optimizer = "bobyqa"),
                nAGQ = 1)

##semantics
semh4 = glmer(Recall ~ (1|Partno) + SCOS + SLSA + SFSG + SIntercept,
                data = hyp4data,
                family = binomial,
                control = glmerControl(optimizer = "bobyqa"),
                nAGQ = 1)

##thematics
themeh4 = glmer(Recall ~ (1|Partno) + TCOS + TLSA + TFSG + TIntercept,
                data = hyp4data,
                family = binomial,
                control = glmerControl(optimizer = "bobyqa"),
                nAGQ = 1)

####table set up####
##create a blank table
tableprint = matrix(NA, nrow = 15, ncol = 5)
colnames(tableprint) = c("Variable", "$b$", "$SE$", "$z$", "$p$")

tableprint[1:5, 2:5] = coef(summary(assoch4))
tableprint[1:5, 1] = dimnames(coef(summary(assoch4)))[[1]]

tableprint[6:10, 2:5] = coef(summary(semh4))
tableprint[6:10, 1] = dimnames(coef(summary(semh4)))[[1]]

tableprint[11:15, 2:5] = coef(summary(themeh4))
tableprint[11:15, 1] = dimnames(coef(summary(themeh4)))[[1]]

tableprint[ , 2] = apa(as.numeric(tableprint[ , 2]),3,T) 
tableprint[ , 3] = apa(as.numeric(tableprint[ , 3]),3,T)
tableprint[ , 4] = apa(as.numeric(tableprint[ , 4]),3,T)
tableprint[ , 5] = apply(as.data.frame(as.numeric(tableprint[ , 5])), 1, p.value)

apa_table(as.data.frame(tableprint),
          align = c("l", rep("c", 4)), 
          caption = "MLM Statistics for Hypothesis 4",
          note = "Each judgment-database bias and sensitivity predicting recall for corresponding judgment block. A: Associative, S: Semantic, T: Thematic.",
          escape = FALSE
 )

```

In our fourth and final hypothesis, we investigated whether the JOR slopes and intercepts obtained in Hypothesis 1 would be predictive of recall ability. Whereas Hypothesis 3 indicated that word relatedness was directly related to recall performance, this hypothesis instead looked at whether or not participants' sensitivity and bias to word relatedness could be used a predictor of recall [@Maki2007]. This analysis was conducted with a multilevel logistic regression, as described in Hypothesis 3, where each database slope and intercept was used as predictors of recall using participant as a random intercept factor. These analyses were separated by judgment condition, so that each set of JOR slopes and intercepts were used to predict recall. The separation controlled for the number of variables in the equation, as all slopes and intercepts would have resulted in overfitting. These values were obtained from Hypothesis 1 where each participant's individual slopes and intercepts were calculated for associative, semantic, and thematic JOR conditions. Table \@ref(tab:hyp1-table1) shows average slopes and intercepts for recall for each of the three types of memory, and Table \@ref(tab:hyp4-table) portrays the regression coefficients and statistics. In the associative condition, FSG slope significantly predicted recall (*b* = `r tableprint[4, 2]`, *p* = `r tableprint[4, 5]`), while COS slope (*b* = `r tableprint[2, 2]`, *p* = `r tableprint[2, 5]`) and LSA slope (*b* = `r tableprint[3, 2]`, *p* = `r tableprint[3, 5]`) were non-significant. In the semantic condition, COS slope (*b* = `r tableprint[7, 2]`, *p* `r tableprint[7, 5]`) and LSA slope (*b* = `r tableprint[8, 2]`, *p* = `r tableprint[8, 5]`) were both found to be significant predictors of recall. FSG slope was non-significant in this condition (*b* = `r tableprint[9, 2]`, *p* = `r tableprint[9, 5]`). Finally, no predictors were significant in the thematic condition, though LSA slope was found to be the strongest (*b* = `r tableprint[13, 2]`, *p* = `r tableprint[13, 5]`). This analysis indicated the extent to which the cognitive processes are related to each other as part of the memory network (i.e., judgment sensitivity predicting recall), furthering the previous two analyses, which illustrated the nature of those cognitive processes' relationship with the underlying memory network. 

# Discussion

This study investigated the relationship between associative, semantic, and thematic word relations and their effect on participant JORs and recall performance through the testing of four hypotheses. In our first hypothesis, we show that bias and sensitivity findings first proposed by @Maki2007a successfully replicated in the associative condition, with slope and intercept values falling within the expected range. While these findings were not fully replicated when extending the analysis to include semantic and thematic JORs (as slopes in these conditions did not fall within the appropriate range), participants still displayed high intercepts and shallow slopes, suggesting overconfidence in judgment making and an insensitivity to changes in strength between pairs. Additionally, when looking at the frequency that each predictor was the strongest in making JORs, FSG was the best predictor for both the associative and semantic conditions, while LSA was the best predictor in the thematic condition. In each of the three conditions, COS was the weakest predictor, even when participants were asked to make semantic judgments. This finding suggests that associative relationships seem to take precedence over semantic relationships when judging pair relatedness, regardless of what type of JOR is being elicited. Additionally, this finding may be taken as further evidence of a separation between associative information and semantic information, in which associative information is always processed, while semantic information may be suppressed due to task demands [@Buchanan2010; @Hutchison2007]. 

Our second hypothesis examined the three-way interaction between FSG, COS, and LSA when predicting participant JORs. At low semantic overlap, a seesaw effect was found in which increases in thematic strength led to decreases in associative predictiveness. This finding was then replicated in Hypothesis 3 when extending the analysis to predict recall. By limiting the semantic relationships between pairs, an increased importance is placed on the role of associations and thematics when making relatedness judgments or retrieving pairs. In such cases, increasing the amount of thematic overlap between pairs results in thematic relationships taking precedent over associative relationships. However, when semantic overlap was high, a complementary relationship was found in which increases in thematic strength in turn led to increases in the strength of FSG as a predictor. This result suggests that at high semantic overlap, associations and thematic relations build upon one another. Because thematics is tied to both semantic overlap and item associations, the presence of strong thematic relationships between pairs during conditions of high semantic overlap boosts the predictive ability of associative word norms for both recall and JORs. 

Finally, our fourth hypothesis used the JOR slopes and intercepts calculated in Hypothesis 1 to investigate whether participants' bias and sensitivity to word relatedness could be used to predict recall. For the associative condition, the FSG slope significantly predicted recall. In the semantic condition, recall was significantly predicted by both the COS and LSA slopes, with COS being the strongest. However, for the thematic condition, although the LSA slope was the strongest, no predictors were significant. One explanation for this finding is that thematic relationships between item pairs act as a blend between associations and semantics. As such, LSA faces increased competition from the associative and semantic database norms when predicting recall in this manner. Additionally, the dominance of FSG when predicting recall in the associative condition may be attributed to word associations being more accessible (and, thus, easier to process) than semantic or thematic relations between pairs.

Overall, our findings indicated the degree to which the processing of associative, semantic, and thematic information impacts retrieval and judgment making tasks and the interactive relationship that exists between these three types of lexical information. While previous research has shown that memory networks are divided into separate systems which handle storage and processing for meaning and association [see @Ferrand2004 for a review], the presence of these interactions suggests that connections exist between these individual memory networks, linking them to one another. As such, we suggest that these memory systems may be connected in such a way as to form a three-tiered, interconnected system. Within this framework, information enters the semantic memory network, which processes features of concepts and provides a means of categorizing items based on the similarity of their features. The associative network adds information for items based on contexts generated by reading or speech. The thematic network then pulls in information based on semantic and associative input to create a mental representation of both the item and its place in the world relative to other concepts.

While this study did not explore the timing or order in which information is input from each of these systems of concept information, it may be comparable to the dual-route model of reading and naming [@Coltheart1993], in that each system runs in parallel when contributing to the judgment and recall process [See also @Coltheart2001]. Alternatively, prevous research on early linguistic access suggests that these three systems may be operating simultaneously. For example, The Language and Situated Simulation Theory (LASS, see @Barsalou2008 for a review) proposes that meaning is derived from the concurrent processing of both linguistic information (e.g., visual stimuli) and modal simulations (i.e., mental imagery related to the stimulus object). According to this theory, when an object is read, both systems become imemediately engaged, though the linguistic system is assumed to be faster, because mental representations of linguistic forms are more accessible in memory relative to simulations (Barsalou et al., 2008, @Thomson1973). As the linguistic system becomes activated, it in turn activates related linguistic forms (i.e., associated words). Because word associations play a key role within this linguistic system, associative information may then come online first relative to semantic concept information.

Viewing this model purely through the lens of semantic memory, it draws comparison to dynamic attractor models [@Jones2015; @Hopfield1982; @McLeod2000]. One of the defining features of dynamic attractor models is that they allow for some type of bidirectionally or feedback between connections in the network. In the study of semantic memory, these models are useful for taking into account multiple restraints such as links between semantics and the orthography of the concept in question. Our hypothesis extends this notion as a means of framing how these three memory systems are connected. The underlying meaning of a concept is linked with both information pertaining to its co-occurrences in everyday language and information relating to the general contexts in which it typically appears.

How then does this hypothesis lend itself towards the broader context of psycholinguistic research? One application of this hypothesis may be found when examining models of word recognition. One popular class of word recognition models are those based upon the "triangle model" first proposed by @Seidenberg1989a [see @Harley2008 for a review]. The key feature of these models is that they recognize speech and reading based upon the orthography, phonology, and meaning of words in a bidirectional manner, similar to the models described above. @Harm2004 developed a version which included a focus on semantics, with word meaning being related to input from the orthography and phonology components of the model. As our findings from the present study suggest that thematic and associative knowledge is integrated with meaning, one way of framing our results within this literature is to consider the semantic section of the triangle model as being comprised of this interconnected system,with concept information processed to some degree on each of these domains. Thus, one area for future studies of this nature may be investigating how aspects of orthography and phonology impact these memory networks. 

Finally, future studies may wish to consider elements of thematic and associative knowledge when examining semantic based tasks (such as word recognition and reading), as thematic and associative information is interconnected with the semantic network. To fully understand the interplay of associatitive, semantic, and thematic networks, it will be necessary to control for several types of word properties (e.g., item frequency, concreteness, etc.), as these properties have been shown to influence judgments and recall. Ultimately, our findings show that these three lexical networks are highly related suggest they are individual components of an interconnected system.

## Compliance with Ethical Standards

The authors declare that they have no conflict of interest. The study was approved by the Institutional Review Board at Missouri State University. Particiants filled out an informed consent at the beginning of the study, after accepting the HIT on Mechanical Turk. The complete study with consent form can be found on our OSF page: http://osf.io/y8h7v.

\newpage

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
